### 2022-12-17 Rmd/ICES advice and assessment summaries.Rmd
Restructured into Rmd folder

# s   <- "hke-nrtn"
# 
# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
# 
# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
# 
# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
# 
# Stock size is well above MSY Btrigger and fishing mortality is below Fmsy. In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years. 
# 
# [hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hke.27.3a46-8abd.pdf)

@@

@@

# s   <- "cod-347d"
# 
# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
# 
# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
# 
# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
# 
# The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years. According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey. The 2017 SSB is 32% lower compared to the estimate that was presented last year. This also leads to a substantially lower advice for 2019 compared to the advice for 2018. 
# 
# [cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.47d20.pdf)

@@

@@

# s   <- "ple-nsea"
# 
# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
# 
# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
# 
# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
# 
# Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy. An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016). This plan is not adopted by Norway, thus, not used as the basis of the advice for this shared stock. The advice is based on the MSY approach. 
# 
# [ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/ple.27.420.pdf)

@@

@@

# s   <- "sol-nsea"
# 
# cat(filter(iAdvice, stockkeylabelold == s, assessmentyear == lay)$stocklongname)
# 
# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment")) 
# 
# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced=TRUE)
# 
# Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea. 
# 
# [sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/sol.27.4.pdf)

@@

>>>>

s   <- "hke-nrtn"

@@

Stock size is well above MSY Btrigger and fishing mortality is below Fmsy. In the past, there has been a tendency to underestimate SSB and overestimate F but this is not the case in the most recent years.

[hke.27.3a46-8abd.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/hke.27.3a46-8abd.pdf)

@@

s   <- "cod-347d"

@@

The 2018 assessment gave substantially lower estimates of SSB and higher estimates of fishing mortality in recent years. According to the advice document, this is (partially) caused by a lower than expected survey index for the strong 2013 and 2016 year classes in IBTS quarter 1 survey and the IBTS quarter 3 survey. The 2017 SSB is 32% lower compared to the estimate that was presented last year. This also leads to a substantially lower advice for 2019 compared to the advice for 2018.

[cod.27.47d20.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.47d20.pdf)

@@

s   <- "ple-nsea"

@@

Stock size is well above MSY Btrigger (and close to 1 million tonnes) and fishing mortality is around Fmsy. An EU multiannual management plan (MAP) has been proposed for this stock (EU, 2016). This plan is not adopted by Norway, thus, not used as the basis of the advice for this shared stock. The advice is based on the MSY approach.

[ple.27.420.pdf](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/ple.27.420.pdf)

@@

s   <- "sol-nsea"

@@

Stock size above MSY Btrigger, fishing mortality around Fmsy. Advice according to the EU multiannual plan (MAP) for the North Sea.

[sol.27.4.pdf](http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/sol.27.4.pdf)

### 2022-12-17 Rmd/Plot for pelagic forum 2019.Rmd
Restructured into Rmd folder

# -----------------------------------------------------------------------------------------------
# ICES advice and assessment summaries.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# -----------------------------------------------------------------------------------------------

@@

# Libraries
library(rmarkdown) # creating markdown documents
library(tidyverse) # combined package of dplyr, tidyr, ggplot, readr, purrr and tibble
library(reshape2)  # reshaping data; e.g. dcast
library(ggthemes)  # for themes
library(pander)    # for print tables
library(readxl)    # read excel files
library(cowplot)   # multiplots
library(directlabels)  # label graphs
library(RColorBrewer) # colours
library(lubridate)

# Load utils code
source("../../mptools/r/my_utils.r")

dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# Load dataset
load(file=paste(dropboxdir, "/rdata/iAssess.RData",sep=""))
load(file=paste(dropboxdir, "/rdata/iAdvice.RData",sep=""))

fay <- 2011
fy  <- 2000
lay <- 2019
ly  <- 2019

# stock=c("spr-nsea","spr.27.3a4"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark==TRUE
# stock=c("mac-nea"); firstassessyear=2016;lastassessyear=2019;firstyear=2000;lastyear=2020; include.benchmark=TRUE; plot=c("stocksize","fishingpressure","recruitment")

# d %>% filter(stockkeylabelold == "mac-nea", assessmentyear == 2019) %>% View()
# unique(iAssess$purpose)

# Function to plot historical assessments
plot_history <- function(stock,firstassessyear,lastassessyear, firstyear, lastyear, include.benchmark = FALSE, 
                         plot=c("stocksize","fishingpressure")) {

  d <-
    iAssess %>% 
    ungroup() %>%

    {if (include.benchmark == TRUE) {
      filter(., purpose %in% c("advice","benchmark", "interbenchmark", "bench"))
    } else {
      filter(., purpose %in% c("advice")) 
    }} %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%

    filter(assessmentyear   %in% c(firstassessyear:lastassessyear), 
           year             %in% firstyear:lastyear) %>%

    mutate(tyear     = substr(as.character(assessmentyear), 3,4)) %>%

    #set type
    mutate(
      type = ifelse(assessmentyear == lastassessyear & purpose == "advice"       , "last"      , NA),
      type = ifelse(assessmentyear == lastassessyear & grepl("bench",purpose),  "last_bench", type),  
      type = ifelse(assessmentyear != lastassessyear & purpose == "advice"       , "advice"    , type),
      type = ifelse(assessmentyear != lastassessyear & grepl("bench",purpose),  "benchmark" , type),
      type = ifelse(assessmentyear != lastassessyear & grepl("replaced",purpose) , "replaced"  , type),
      type = ifelse(is.na(type), "other", type)) %>%

    # set group
    mutate(group = paste0(tyear, type)) %>%

    select(assessmentyear, tyear, year, stockkey, stockkeylabel, stockkeylabelold, stockkeylabelnew,
           type, group, source,
           lowrecruitment, recruitment, highrecruitment, 
           lowfishingpressure, fishingpressure, highfishingpressure, 
           lowstocksize, stocksize, highstocksize) %>%
    data.frame()

  # find out what is the last assessment
  last <- filter(d, assessmentyear == lastassessyear) %>% distinct(type) %>% as.character()

  # plot stocksize
  stocksize <-
    d %>% 
    filter(!is.na(stocksize)) %>%

    ggplot(aes(year,stocksize, group=group)) +

    theme_publication() +

    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          plot.title  = element_text(size=10), 
          plot.margin = margin(2, 2, 2, 2, "mm"),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(data=filter(d, type==last), 
                aes(x=year, ymin=lowstocksize, ymax=highstocksize, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +

    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=blim, label="blim", vjust=-1), 
               colour="red", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
               aes( firstyear, blim, label = "blim", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes = FALSE) +

    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=msybtrigger), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
               aes( firstyear, msybtrigger, label = "msy btrigger", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "stocksize")

  # fishingpressure
  fishingpressure <-
    d %>% 
    filter(!is.na(fishingpressure)) %>%

    ggplot(aes(year,fishingpressure, group=group)) +

    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          plot.title  = element_text(size=10), 
          plot.margin = margin(2, 2, 2, 2, "mm"),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(data=filter(d, type==last), 
                aes(x=year, ymin=lowfishingpressure, ymax=highfishingpressure, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +

    geom_hline(data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
               aes(x=assessmentyear, yintercept=fmsy), 
               colour="red", linetype="dashed", inherit.aes=FALSE) +
    geom_text (data=filter(iAdvice, stockkeylabelold %in% stock | stockkeylabel %in% stock, assessmentyear == lastassessyear),
               aes( firstyear, fmsy, label = "fmsy", vjust = -0.3, hjust = 0),
               colour="red", inherit.aes=FALSE) +

    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "fishingpressure")

  # Recruitment
  recruitment <-
    d %>% 
    filter(!is.na(recruitment)) %>%

    ggplot(aes(year,recruitment, group=group)) +

    theme_publication() +
    theme(legend.title=element_blank(),
          axis.text.x = element_text(angle = 0, vjust = 0.5, size=9),
          axis.text.y = element_text(size=9),
          plot.title  = element_text(size=10), 
          plot.margin = margin(2, 2, 2, 2, "mm"),
          # strip.background = element_blank(),
          legend.position = "null") +

    geom_ribbon(data=filter(d, type==last), 
                aes(x=year, ymin=lowrecruitment, ymax=highrecruitment, fill=type), 
                alpha=0.2) +
    geom_line(aes(colour = type, size=type, linetype=type) ) +

    geom_dl(aes(label  = tyear, colour = type), method = list(dl.combine("last.points"), cex = 0.8)) +

    scale_colour_manual  (values=c(last ="red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
    scale_fill_manual    (values=c(last = "red",last_bench="blue",advice="black",benchmark="blue",replaced="gray40",other="gray80")) +
    scale_linetype_manual(values=c(last="solid",last_bench="solid",advice="solid",benchmark="solid",replaced="solid",other="solid")) +
    scale_size_manual    (values=c(last= 1.2,last_bench=1.2,advice=0.8,benchmark=0.8,replaced=0.8,other=0.8)) +

    expand_limits(y = 0) +
    xlim(firstyear,lastyear) +
    labs(x = NULL, y = NULL , title = "Recruitment")

    if (length(plot) == 1) {
      if ("stocksize" %in% plot ) print(stocksize)
      if ("fishingpressure" %in% plot ) print(fishingpressure)
      if ("recruitment" %in% plot ) print(recruitment)

    } else if (length(plot) == 2) {

      if ("stocksize" %in% plot & "fishingpressure" %in% plot) plot_grid(stocksize, fishingpressure,  ncol=2, align = 'hv') 
      if ("stocksize" %in% plot & "recruitment" %in% plot) plot_grid(stocksize, recruitment,  ncol=2, align = 'hv') 
      if ("recruitment" %in% plot & "fishingpressure" %in% plot) plot_grid(fishingpressure, recruitment,  ncol=2, align = 'hv')

    } else if (length(plot) == 3) {
        plot_grid(stocksize, fishingpressure,  recruitment, ncol=length(plot), align = 'hv') 
    }

  } # End of plot_history function

# plot_history(stock=s,firstassessyear=fay,lastassessyear=lay, firstyear=fy, lastyear=ly, include.benchmark = TRUE,               plot = c("stocksize", "fishingpressure", "recruitment"), include.benchmark = TRUE)

# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("spr-nsea","spr.27.3a4"); include.replaced <- TRUE
# firstadviceyear <- 2016; lastadviceyear  <- 2020; stock<-c("mac-nea"); include.replaced <- TRUE

# Function to plot advice overviews
table_advice <- function(stock,firstadviceyear,lastadviceyear, include.replaced=FALSE) {

  d <-
    iAdvice %>%

    filter(stockkeylabelold %in% stock | stockkeylabel %in% stock) %>%

    filter(tacyear %in% firstadviceyear:lastadviceyear) %>%

    {if (include.replaced == TRUE) {
      filter(., purpose %in% c("advice","replaced"))
    } else {
      filter(., purpose %in% c("advice")) 
    }} %>%

    mutate(advicebasis= substr(advicebasis,1,25),

           advice     = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicemax  = ifelse(!is.na(advisedcatchmax), advisedcatchmax, advisedlandingsmax),
           advicetype = ifelse(!is.na(advisedcatchmax), "catch", "landings"),

           tac        = ifelse(!is.na(tac), tac, tal),
           tactype    = ifelse(!is.na(tac), "catch", "landings"),

           catch      = ifelse(!is.na(catches), catches, NA),
           catchtype  = ifelse(!is.na(catches), "catches", ""),

           catchtype  = ifelse(!is.na(landings) & is.na(catch), "landings", catchtype),
           catch      = ifelse(!is.na(landings) & is.na(catch), landings, catch),

           catchtype  = ifelse(!is.na(officiallandings) & is.na(catch), "officiallandings", catchtype),
           catch      = ifelse(!is.na(officiallandings) & is.na(catch), officiallandings, catch),

           adv_c      = 100 * (advicemax / lag(advicemax, n=1) - 1),
           tac_c      = 100 * (tac /lag(tac, n=1) - 1),
           cat_c       = 100 * (catch / lag(catch, n=1) -1)
           ) %>%

    select(stockkeylabel, tacyear, advicebasis, purpose, advice, fadvmax, advicemax, tac, catch, 
           adv_c, tac_c, cat_c,
           advicetype, tactype, catchtype)

  d %>% 
  select("stockkeylabel", "tacyear", "advicebasis", "purpose", "fadvmax", "advicemax", "tac", "catch") %>%
  arrange(tacyear, desc(purpose)) %>% 
  pandoc.table(., 
               style        = "simple",
               split.tables = 200, 
               justify      = "right",
               missing      =" ",
               big.mark     = '', 
               round        = c(0,0,0,2,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
}

# table_advice(stock=s,firstadviceyear=fay+1,lastadviceyear=lay+1, include.replaced = TRUE)

@@

**Summary of ICES 2019 assessments and advice**

@@

In this document an overview is presented of the ICES assessment and advice for the most recent number of years for the stocks that are relevant to the PFA and Dutch fisheries in general.

<!--Section 2: material and methods ----------------------------------------------------------- -->

# Material and methods

The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.

<!--Section 3: advice and assessment overviews ------------------------------------------------ -->

# Assessment and advice overviews

>>>>

# -----------------------------------------------------------------------------------------------
# Plots for pelagic forum 2019.Rmd
#
# 06/07/2018 first coding
# 14/12/2018 updated for the new database structure
# 01/06/2019 updated for herring advice round
# 29/09/2019 adapted from stock summary code
# -----------------------------------------------------------------------------------------------

@@

library(tidyverse)
library(cowplot)
library(directlabels)  # for printing labels at end of geom lines
library(scales)
library(stringr)
library(readxl)
library(gridExtra)                   # for plotting table as image (bio data)
library(RColorBrewer)

# Load utils code
source("../../mptools/r/my_utils.r")
source("iPlot_utils.r")

# Set dropbox folder
dropboxdir <- paste(get_dropbox(), "/iAdvice", sep="")

# load the data
iSpecies <-
  get(load(file=paste(dropboxdir,"/rdata/iSpecies.RData", sep=""))) %>% 
  ungroup() %>% 
  dplyr::select(speciesfaocode, trophicguild, fisheriesguild, sizeguild)

iAssess <-
  get(load(file=paste(dropboxdir,"/rdata/iAssess.RData", sep=""))) %>% 
  left_join(iSpecies, by="speciesfaocode")

load(file=paste(dropboxdir,"/rdata/iAdvice.RData", sep=""))
load(file=paste(dropboxdir,"/rdata/iRename.RData", sep=""))
load(file=paste(dropboxdir,"/rdata/iStockkey.RData", sep=""))

# add intermediate and forecast data
forecasts <-
  read_excel(path=paste(dropboxdir,"/excel/Pelagic forecasts 2019.xlsx", sep=""), col_names = TRUE) %>%
  left_join(iStockkey, by="stockkeylabelold") %>%
  mutate(stockkeylabel = stockkeylabelnew) %>%
  left_join(iRename,   by=c("stockkey", "stockkeylabel")) %>%
  left_join(iSpecies,  by="speciesfaocode") %>%
  mutate(catcheslandingsunits = "tonnes")

# t <-
#   icesSAGfull %>%
#   filter(year < assessmentyear) %>% 
#   bind_rows(forecasts)

fay <- 2011
fy  <- 1990
lay <- 2019
ly  <- 2019

@@

**Status of pelagic stocks in the Northeast Atlantic**

@@

In this document an overview is presented of the ICES assessment and advice for pelagic stocks in the Northeast Atlantic. The overviews are based on two databases that have been developed as part of a more comprehensive research effort into the role of fisheries advice in European decision-making: one database containing the history of stock assessments and one database containing the history of scientific advice and management decisions.

<!--Section 2: general trends ------------------------------------------------ -->

# General trends

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

# catches by fisheriesguild

d <-
  iAssess %>% 
  ungroup() %>% 
  # filter(assessmentyear      == lay | (assessmentyear == 2017 & stockkeylabelold == "cap-icel") ) %>% 
  filter(assessmentyear      == lay  ) %>% 
  # filter(catcheslandingsunits == "tonnes") %>% 
  filter(year >= fy, year < ly) %>% 
  filter(purpose %in% c("advice")) %>%

  mutate(catches = ifelse(is.na(catches) | catches == 0, landings, catches))  %>% 
  mutate(fisheriesguild = ifelse(fisheriesguild %in% c("pelagic","demersal"), fisheriesguild, "other"))  %>% 
  mutate(fisheriesguild = factor(fisheriesguild, levels=c("pelagic","demersal","other"))) %>%

  group_by(fisheriesguild, speciesfaocode, year) %>% 
  summarize(catches  = sum(catches, na.rm=TRUE),
            landings = sum(landings, na.rm=TRUE)) %>%

  group_by(fisheriesguild, year) %>% 
  summarize(catches  = sum(catches, na.rm=TRUE),
            landings = sum(landings, na.rm=TRUE)) %>% 
  group_by(year) %>% 
  mutate(perc = round(100 * catches/ sum(catches), digits=2)) %>% 
  arrange(catches) %>% 
  mutate(fisheriesguild = factor(fisheriesguild, fisheriesguild))

# d %>% filter(year == 2018) %>% arrange(-catches) %>% View()

d2 <-
  d %>% 
  group_by(fisheriesguild) %>% 
  summarize(perc = mean(perc),
            catches=mean(catches),
            year = mean(year)) %>% 
  ungroup() %>% 
  mutate(cumcatches=cumsum(catches))

# plot stacked catches: pelagic, demersal and other
p1 <-
  d %>%

  ggplot(aes(year,catches)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_bar(aes(fill=(fisheriesguild)), stat="identity", position="stack", colour="black") +
  # geom_text(data=subset(d, year==ly-1), aes(label=round(perc,0), group=fisheriesguild), 
  #           position = position_stack(vjust = 0.5), colour="white") +
  geom_text(data=subset(d, year==ly-1), 
            aes(label=fisheriesguild, group=fisheriesguild, x=ly), 
            position = position_stack(vjust = 0.5), colour="black", hjust=0) +
  # geom_dl(aes(label= perc),
  #         position = position_stack(vjust = 0.5),
  #         method   = list(dl.trans(x=x+0.9), dl.combine("last.points"),
  #                         fontface = 'bold',cex=.9, hjust = 0)) +
  scale_fill_brewer(palette = "Paired") +
  labs(x="",y="tonnes",title="catches") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly+3))

# d %>% filter(year == 2018) %>% View()

# plot of filled relative catches (how to add the direct labels?)
p2 <-
  d %>% 
  ggplot(aes(year,perc)) +
  theme_publication() +
  theme(legend.position="none") +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  theme(plot.background = element_blank()) +
  theme(axis.text = element_blank()) +
  theme(axis.ticks = element_blank()) +
  geom_bar(aes(fill=fisheriesguild), stat="identity", position="stack", colour="black") +
  scale_fill_brewer(palette = "Paired") +
  labs(x="",y="",title="") +
  # scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly))

vp <- grid::viewport(width = 0.3, height = 0.3, 
                     x = 0.15, y = 0.12, 
                     just = c("left", "bottom"))

 print(p1)
     theme_set(theme_bw(base_size = 8))
     # theme_white()
     print(p2, vp = vp)
     theme_set(theme_bw())

  # cowplot::plot_grid(p1, p2,
  #                ncol=2, scale=c(1.0, 0.25), align="hv",
  #                rel_widths = c(1.0, 1.0), rel_heights = c(1.0, 1.0))

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

# catches of pelagic species

d <-
  iAssess %>% 
  ungroup() %>% 
  filter(year >= fy, year < ly) %>% 
  filter(purpose %in% c("advice")) %>% 
  filter(fisheriesguild == "pelagic") %>%
  filter(assessmentyear == lay ) %>%

  mutate(speciesfaocode = ifelse(speciesfaocode %in% c("mac","whb","her", "hom", "pil","spr", "cap"), 
                                 speciesfaocode, "other")) %>%

  # filter(speciesfaocode=="ane") %>% View()
  # filter((speciesfaocode == "ane.27.8" & assessmentyear == lay-1) | 
  #        (speciesfaocode != "ane.27.8") & assessmentyear == lay ) %>%

  mutate(
    catches = ifelse(!is.na(catches), catches, landings) ) %>% 
  dplyr::select(-landings) %>% 
  group_by(speciesfaocode, year) %>% 
  summarize(
    catches  = sum(catches, na.rm=TRUE)) %>% 
  group_by(speciesfaocode) %>% 
  mutate(
    totalcatches = sum(catches, na.rm=TRUE) ) %>%
  filter(totalcatches > 0) %>% 
  group_by(year) %>% 
  mutate(
    perc = round(100 * catches/ sum(catches), digits=2) ) %>% 
  arrange(catches) %>% 
  mutate(speciesfaocode = factor(speciesfaocode, speciesfaocode))

# d %>% filter(speciesfaocode=="ane.27.8") %>%  View()
# d %>% filter(year == 2018) %>% View()

d2 <-
  d %>% 
  group_by(speciesfaocode) %>% 
  summarize(perc = mean(perc),
            catches=mean(catches),
            year = mean(year)) %>% 
  ungroup() %>% 
  mutate(cumcatches=cumsum(catches))

# plot stacked catches: pelagic, demersal and other
p1 <-
  d %>%

  ggplot(aes(year,catches)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_bar(aes(fill=(speciesfaocode)), stat="identity", position="stack", colour="black") +
  geom_text(data=subset(d, year==ly-1), 
            aes(x = year+0.7,label=speciesfaocode, group=speciesfaocode, colour=speciesfaocode ), 
            position = position_stack(vjust = 0.5), hjust=0) +
  # geom_dl(aes(label= perc),
  #         position = position_stack(vjust = 0.5),
  #         method   = list(dl.trans(x=x+0.9), dl.combine("last.points"),
  #                         fontface = 'bold',cex=.9, hjust = 0)) +
  scale_fill_brewer(palette = "Paired") +
  scale_colour_brewer(palette = "Paired") +
  labs(x="",y="tonnes",title="catches by species") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly+1))

d %>% filter(year==2018) %>% ungroup() %>% summarize(catches=sum(catches))

# plot of filled relative catches (how to add the direct labels?)
p2 <-
  d %>% 
  ggplot(aes(year,perc)) +
  theme_publication() +
  theme(legend.position="none") +
  theme(plot.margin = unit(c(0,0,0,0), "cm")) +
  theme(plot.background = element_blank()) +
  theme(axis.text = element_blank()) +
  theme(axis.ticks = element_blank()) +
  geom_bar(aes(fill=speciesfaocode), stat="identity", position="stack", colour="black") +
  scale_fill_brewer(palette = "Paired") +
  labs(x="",y="",title="") +
  # scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly))

vp <- grid::viewport(width = 0.3, height = 0.3, 
                     x = 0.15, y = 0.12, 
                     just = c("left", "bottom"))

 print(p1)
     theme_set(theme_bw(base_size = 8))
     # theme_white()
     print(p2, vp = vp)
     theme_set(theme_bw())

  # cowplot::plot_grid(p1, p2,
  #                ncol=2, scale=c(1.0, 0.25), align="hv",
  #                rel_widths = c(1.0, 1.0), rel_heights = c(1.0, 1.0))

```{r, echo=FALSE, fig.asp=0.7, fig.align="center", message=FALSE, warning=FALSE}

# catches and advice for key pelagic species
d <-
  iAssess %>%
  filter(year < ly) %>% 
  bind_rows(forecasts) %>% 
  filter(stockkeylabelold %in% c("her-47d3","her-noss","mac-nea","whb-comb")) %>% 
  filter(assessmentyear == lay ) %>%
  filter(year >= fy, year <= ly+1) %>% 
  filter(purpose %in% c("advice")) %>%

  mutate(
    catches = ifelse(!is.na(catches), catches, landings) ) %>% 
  dplyr::select(-landings) %>% 
  group_by(stockkeylabelold, year) %>% 
  summarize(
    catches  = sum(catches, na.rm=TRUE)) %>% 
  group_by(stockkeylabelold) %>% 
  mutate(
    totalcatches = sum(catches, na.rm=TRUE) ) %>%
  filter(totalcatches > 0) %>% 
  group_by(year) %>% 
  mutate(
    perc = round(100 * catches/ sum(catches), digits=2) ) %>% 
  arrange(catches) %>% 
  mutate(stockkeylabel = factor(stockkeylabelold, stockkeylabelold)) %>% 
  arrange(stockkeylabel, year)

# plot stacked catches: pelagic, demersal and other
d %>% 
  ggplot(aes(year,catches)) +
  theme_publication() +
  theme(legend.position="none") +
  geom_bar(aes(fill=(stockkeylabelold)), stat="identity", position="stack", colour="black") +
  geom_text(data=subset(d, year==ly+1), 
            aes(x = year+0.7,label=stockkeylabelold, group=stockkeylabelold, colour=stockkeylabelold ), 
            position = position_stack(vjust = 0.5), hjust=0) +
  scale_fill_brewer(palette = "Paired") +
  scale_colour_brewer(palette = "Paired") +
  labs(x="",y="tonnes",title="catches by key pelagic species") +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(limits=c(fy,ly+5))

# d %>% 
#   arrange(stockkeylabelold, year) %>% 
#   filter(year >= 2019) %>% 
#   group_by(year) %>% 
#   summarize(catches=sum(catches)) %>% 
#   View()

<!--Section 3: overviews by stock ------------------------------------------------ -->

# Overviews by stock
